# Prior Knowledge Integration

```{r "setup", include=FALSE}
library(reticulate)
knitr::knit_engines$set(python = reticulate::eng_python)
```

`ariadne` (R) and `ariadnepy` (Python) give access to the same resource: a
single graph assembled from public biological databases (Gene Ontology,
KEGG, UniProt, BugSigDB, ChocoPhlAn, and more), which can be queried and
traversed to link one type of biological identifier to another — for
example, taxonomic names to functional pathways, or gene families to
disease modules.

This page walks through the same short analysis twice, once in R and once
in Python, so you can follow along in whichever language you use.

## Installation

::: {.panel-tabset}

## R

```{r "install-r"}
#| eval: false
remotes::install_github("Minotau-R/ariadne")
```

## Python

```{python "install-py"}
#| eval: false
pip install ariadnepy
```

:::

## Building the resource graph

Both packages expose a single entry point that downloads (and locally
caches) the resource graph.

::: {.panel-tabset}

## R

```{r "build-graph-r"}
library(ariadne)

graph <- ariadne()
summary(graph)
```

## Python

```{python "build-graph-py"}
import ariadnepy

graph = ariadnepy.ariadne()
print(graph.summary())
```

:::

## Searching for paths

Before linking two resources, it helps to see which routes connect them.
`searchPath()` / `search_path()` print the first `k` shortest paths.

::: {.panel-tabset}

## R

```{r "search-path-r"}
searchPath(graph, ko ~ ec, k = 3)
```

## Python

```{python "search-path-py"}
ariadnepy.search_path(graph, "ko ~ ec", k=3)
```

:::

## Visualising a path

`plotPath()` / `plot_path()` draw the full resource graph, highlighting a
chosen path in red.

::: {.panel-tabset}

## R

```{r "plot-path-r"}
plotPath(graph, ko ~ ec, k = 1, prune = TRUE, focus = TRUE)
```

## Python

```{python "plot-path-py"}
fig = ariadnepy.plot_path(graph, "ko ~ ec", k=1, prune=True, focus=True)
fig
```

:::

## Weaving a linkmap

`weavePath()` / `weave_path()` traverse the chosen path and return a
two-column table mapping origin identifiers to target identifiers.

::: {.panel-tabset}

## R

```{r "weave-path-r"}
ko2ec <- weavePath(graph, ko ~ ec, k = 1)
head(ko2ec)
```

## Python

```{python "weave-path-py"}
ko2ec = ariadnepy.weave_path(graph, "ko ~ ec", k=1)
ko2ec.head()
```

:::

`weavePath()` / `weave_path()` also accept a fixed path table (e.g. from
`drawPath()` / `draw_path()`, see below) instead of the full graph, which
skips path search entirely and re-runs just those steps — useful once
you've settled on a route and want to reuse it with different `init` IDs.

## Reproducibility: recording the path used

`drawPath()` / `draw_path()` return the resource, version, and URL used at
every hop of a path — handy for citing the exact data sources behind a
result.

::: {.panel-tabset}

## R

```{r "draw-path-r"}
drawPath(graph, ko ~ ec, k = 1)
```

## Python

```{python "draw-path-py"}
ariadnepy.draw_path(graph, "ko ~ ec", k=1)
```

:::

## Python only: annotating AnnData objects

`ariadnepy` additionally provides `add_modules()` / `get_modules()`, which
attach a linkmap as module-membership columns directly onto an `AnnData`
object's `var` (features) or `obs` (samples) metadata — no R equivalent is
needed here since `AnnData` is Python-specific. Any two-column linkmap
works, whether it comes from `weave_path()`/`weave_complex()` or from
elsewhere; the example below builds one by hand from the bundled
butyrate-producer dataset, so it runs without any network access.

```{python "anndata-py"}
import anndata as ad
import numpy as np

butyrate = ariadnepy.load_butyrate()

adata = ad.AnnData(
    X=np.zeros((1, len(butyrate))),
    var=butyrate.set_index("taxname"),
)

# A minimal linkmap: taxon -> a butyrate-production module ID.
linkmap = butyrate[["taxname"]].assign(module="M_BUTYRATE")

adata = ariadnepy.add_modules(adata, linkmap, key="index")
adata.var.head()
```

## Session info

::: {.callout-note collapse="true"}

## Click to expand 👇

::: {.panel-tabset}

## R

```{r "session-info-r"}
#| cache: false
sessioninfo::session_info(installed.packages()[, "Package"], include_base = TRUE)
```

## Python

```{python "session-info-py"}
import sys
print(sys.version)
print("ariadnepy", ariadnepy.__version__)
```

:::

:::